---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Plotting data in dataframes

```{r}
library(Stat2Data)
data(SeaIce)
```


Most data in R are organized into dataframes.  Similar to when we plot data in a spreadsheet, to plot data from a dataframe we need to tell R exactly what we want on the **x-axis (horizontal)** and **y-axis (vertical)**.  


`r emo::ji("warning")` **Note:** For reasons we don't have to get in to, the `lynx` data were a special case where we didn't have to define x and y`r emo::ji("warning")`

We can plot the Extent of artic sea ice again using the `plot()` command, and using a cool convention in R called **formula notation**.  Formula notation uses the (**tilde**)[https://en.wikipedia.org/wiki/Tilde] symbol `~`.  In math, `~` can have several meanings.  In R, it means "relates to" , "versus", "depends on."  So we can plot the relation between `Year` and seac `Extent` as a **y versus x** relation as `Extent ~ Year`.  We also have to include the argument `data = SeaIce` so R knows where to get `Extent` and `Year`.

```{r}
plot(Extent ~ Year, data = SeaIce)  # Note, both words capitalized.
```


## Base-R graphics

When we use the `plot` command were using **Base R** graphics.  As noted before there are several ways to make plots in R and you should be able to spot which one is which when looking at code.  We'll cover some of the key features of Bare R graphics now.  While different plotting methods have different commands and arguments, they all share a common feature: everything in a plot can be customized, and each element is customized with a command or arguement.

### Type of points

`plot()` can draw dots or lines  We make it use lines using the `type = "l"` argument (note that the l is in quotes)

```{r}
plot(Extent ~ Year, type = "l", data = SeaIce)  # Note: l in quotes
```

As noted before, R doesn't mind if you split things on lines.  To keep track of the things I"m doing to the plot I'll format things like this

```{r, eval = F}
plot(Extent ~ Year,  # relationship
     type = "l",     # type of plot; Note: l in quotes
     data = SeaIce)  # data
```

As we did with the `lynx` data we can combine points and lines with `type = "b"`.  (Do you recall what "b" stands for?)
```{r, eval = F}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     data = SeaIce)  # data
```

We can adjust color with `col = ...`.  Recall the this is just a number, not in quotes.
```{r, eval = F}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     data = SeaIce)  # data
```

We can add a main title to with `main = ...`

<!-- Good quiz question: what has to be quote, what doesn't; fix this code but un-qutoing things that should be quoted. -->

```{r}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     data = SeaIce)  # data
```

Its always good to include your units.  `Extent` and `Area` are in square kilometers.  We can say specifically what we want for the y-axis label using `ylab = ...`

```{r}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     ylab = "Extent (square kilometers)",
     data = SeaIce)  # data
```


We can change the appearance of the line using `lty = ...`, which stands for "line type":
```{r}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     ylab = "Extent (square kilometers)",
     lty = 2,        # line type; not quoted
     data = SeaIce)  # data
```


I'm not a fan of the open circles for plotting points; we can change those too using the argument `pch = `.

```{r}
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     ylab = "Extent (square kilometers)",
     lty = 2,        # line type; not quoted
     pch = 16,       # point type; not quoted
     data = SeaIce)  # data
```


### Plotting two columns of data

Often we want to represent two distinct things on our graph.  In spreadsheets these are called separate **series** of data.  When making a **time series plot** like this one we can add a new column of data using a species command called `points()` which works very similar to `plot()`.

`r emo::ji("warning")` **Note:** The `points()` command only works if its it precede by a statement from the `plot()` command `r emo::ji("warning")`

<!-- Simple assignment: fix bad code with points() preceding plot() -->



```{r}
# Main plot: Extent
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     ylab = "Extent (square kilometers)",
     lty = 2,        # line type; not quoted
     pch = 16,       # point type; not quoted
     data = SeaIce)  # data

# Second column of data: Area
points(Area ~ Year, data = SeaIce)
```

I can now customize the sea ice `Area` part of the plot by add **arguements** to the `points()` statement.

```{r}
# Main plot: Extent
plot(Extent ~ Year,  # relationship
     type = "b",     # type of plot; Note: b in quotes
     col = 2,        # color; no quotes
     main = "Arctic Sea Ice Extent", # main title, in quotes
     ylab = "Extent (square kilometers)",
     lty = 2,        # line type; not quoted
     pch = 16,       # point type; not quoted
     data = SeaIce)  # data

# Second column of data: Area
points(Area ~ Year,
       type = "b",
       col = 3,
       pch = 17,
       data = SeaIce)
```

`r emo::ji("question")` There's a problem with my graph though - can you spot it?  It only really becomes apparent when you connect the dots with a line. `r emo::ji("question")`


### Changing the range of a plot axis

Let's go back to our original plot and forget about all the fancy arguments and adding `points()` for a little bit.  The problem with the last graph we made is that some points are not showing - if you look in the lower right-hand part points around 2006-2008 and 2011-2013 are not showing up because the y-axis stops around 3.5.  We can fix this by adding a new argument which sets the limits of the y-axis: `ylim = ...`.  To do this correctly, we have to introduce a new function: `c()`.  This is actually one of the most common functions in R.  In the `c()` we need to tell R the lower and upper limits we want for the y-axis.  Let's do 0 and 8, which will be coded as `c(0,8)`.

So, to set the y-axis limits we do this:

```{r}
plot(Extent ~ Year,  
     ylim = c(0,8), # the c(...) fucntion to set limits
     data = SeaIce) 
```

Now we have a bunch more space at the bottom.  We can add our `points()` back to see if this work:

```{r}
plot(Extent ~ Year,  
     ylim = c(0,8), # the c(...) fucntion to set limits
     data = SeaIce) 
points(Area ~ Year,
      data = SeaIce,
      col = 2)
```


`r emo::ji("warning")` **Note:** `ylim = ...` only goes in `plot()`, not `points()`. `View()` `r emo::ji("warning")`

Let's re-make our fancy plots now with `ylim = ...` set.

```{r}
# Main plot: Extent
plot(Extent ~ Year,  
     type = "b",     
     col = 2,       
     main = "Arctic Sea Ice Extent", 
     ylab = "Extent (square kilometers)",
     lty = 2,        
     pch = 16,       
     ylim = c(0,8), # <#<== y-axis limits
     data = SeaIce)  

# Second column of data: Area
points(Area ~ Year,
       type = "b",
       col = 3,
       pch = 17,
       data = SeaIce)
```


## You try it



### Fixer-uppers

Fix the code below so it works

```{r, eval = F}
plot(Extent , Year,  
     data = SeaIce)  
```

Fix the code below so it works

```{r, eval = F}
plot(Extent ~ Year,  # relationship
     type = b,     # type of plot; Note: b in quotes
     col = "2",        # color; no quotes
     data = SeaIce)  # data
```

### Intermediate

Make a plot with `Area` on the x-axis and `Extent` on the y-axis.

```{r}
## Write the code below
```

<!-- Could add stuff with lynx data if I set it up as a dataframe -->

### Advanced
 
Based on the code above, make a plot of the SeaIce data where `Area` appears within the `plot()` statement, and `Extent` is in `points()`.

```{r}
# Write the code below: 

```



