---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Introduction to ggplot and ggpubr


##SeaIce


```{r}
library(Stat2Data)
library(ggplot2)
library(ggpubr)
```

```{r}
data(SeaIce)
```


```{r}
ggscatter(y = "Area",
          x = "Year",
          data = SeaIce)
```


```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue", #in quotes
          data = SeaIce)
```

```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # not quotes
          data = SeaIce)
```


```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # not quotes
          data = SeaIce)
```

```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)", # in quotes
          data = SeaIce)
```

## Regression line

```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)", 
          add = "reg.line",
          data = SeaIce)
```


```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)", 
          add = "reg.line",
          data = SeaIce) +
  stat_regline_equation()
```



Why is equation different

```{r}
ggscatter(y = "Area",
          x = "t",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)", 
          add = "reg.line",
          data = SeaIce) +
  stat_regline_equation()
```






## Smoother

```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)", 
          add = "loess",
          data = SeaIce)
```


## correlation

```{r}
ggscatter(y = "Area",
          x = "Year",
          color = "blue",
          shape = 16,     # 
          size = 4,       # 
          ylab = "Area (square kilometers)",  
          cor.coef = T,
          data = SeaIce)
```



## Long time series

```{r}
data(CO2Hawaii)
```


Year

Why stacked?
```{r}
ggscatter(y = "CO2",
       x = "Year",
       data = CO2Hawaii)
```

```{r}
ggscatter(y = "CO2",
       x = "Year",
       color = "Month",
       data = CO2Hawaii)
```

What's different?
```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       data = CO2Hawaii)
```


```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       add = "reg.line",
       data = CO2Hawaii)
```


```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       add = "loess",
       data = CO2Hawaii)
```

```


```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       data = CO2Hawaii) +
  xlim(0,100)
```



```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       data = CO2Hawaii) +
  xlim(0,100) +
  ylim(350,365)
```


```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       add = "reg.line",
       data = CO2Hawaii) +
  xlim(0,100) +
  ylim(350,365)
```

```{r}
ggscatter(y = "CO2",
       x = "t",
       color = "Month",
       add = "loess",
       data = CO2Hawaii) +
  xlim(0,100) +
  ylim(350,365)
```



##Moth data - 2 cities

Cook and Grant 2000.  Frequency of insularia during the decline
in melanics in the peppered.  Heredity 85: 580-585.


Make dataframe

##Nottingham data

```{r}
year.nottingham <- seq(from = 1993,
                       to = 2000)
carbonaria.nottingham <- c(58.5, 50.4, 38.5, 27.3, 28.9, 
                           30.9, 19.2, 19)

df.nottingham <- data.frame(year = year.nottingham,
                            N.carbonaria = carbonaria.nottingham,
                            city = "Nottingham")

plot(N.carbonaria ~ year, data =  df.nottingham, type = "b")
```

How many times did R print "Nottingham" How did it make that choice?


```{r}

year.york <- seq(from = 1993,
                 to = 2000)

carbonaria.york <- c( 69.6, 60, 75.0, 51.4, 42.1,
                      41.2, 37.0, 20.8)

df.york <- data.frame(year = year.york,
                      N.carbonaria = carbonaria.york,
                      city = "York")

plot(N.carbonaria ~ year, data =  df.york, type = "b")
```



```{r}
df.moth <- rbind(df.nottingham,
                 df.york)
```



```{r}
is(df.moth)
```


```{r}
dim(df.moth)
```



```{r}
# install.packages("ggplot2")
```


```{r}
library(ggplot2)
library(ggpubr)
```


```{r}
ggscatter(y = "N.carbonaria",   # in quotes
          x = "year",           # in quotes
          data = df.moth)       # in quotes
```


```{r}
ggscatter(y = "N.carbonaria", 
          x = "year",
          color = "city",    # in quotes
          data = df.moth)
```

```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          color = "city",
          shape = "city",     # in quotes
          data = df.moth)
```

```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          color = "city",
          shape = "city",
          size = 5,       # not in quotes!
          data = df.moth)
```



```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          color = "city",
          shape = "city",
          size = 5,       
          data = df.moth,
          ylab = "Number of carbonaria moths", # labels in quotes
          xlab = "Year")   # in quotes
```



## Regression lines


```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          data = df.moth,
          color = "city",
          shape = "city",
          add = "reg.line",
          size = 5)
```

```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          data = df.moth,
          color = "city",
          shape = "city",
          add = "reg.line",
          conf.int = T,
          size = 5)
```



Only 1 line
```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          data = df.moth,
          color = "city",
          shape = "city",
          add = "reg.line",
          conf.int = T,
          size = 5) +
  stat_regline_equation()
```



Facetings - regressions for both lines
```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          data = df.moth,
          color = "city",
          shape = "city",
          add = "reg.line",
          conf.int = T,
          facet.by = "city",
          size = 5) +
  stat_regline_equation()
```



## Correlations


Only 1 correlations coeffient

```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          color = "city",
          shape = "city",
          size = 5,       
          data = df.moth,
          cor.coef = T,
          ylab = "Number of carbonaria moths", # labels in quotes
          xlab = "Year")   # in quotes
```


Faceting gives us 2
```{r}
ggscatter(y = "N.carbonaria",
          x = "year",
          color = "city",
          shape = "city",
          size = 5,       
          data = df.moth,
          cor.coef = T,
          facet.by = "city",
          ylab = "Number of carbonaria moths", # labels in quotes
          xlab = "Year")   # in quotes
```



```{r}
ggplot(data = df.moth,
        aes(y = N.carbonaria,
            x = year,
            color = city)) +
  geom_point() +
  geom_line()
```






## Original data

```{r}
# Cook and Grant 2000.  Frequency of insularia during the decline
# in melanics in the peppered.  Heredity 85: 580-585.
year.nottingham <- seq(from = 1993,
                       to = 2000)
carbonaria.nottingham <- c(58.5, 50.4, 38.5, 27.3, 28.9, 
                           30.9,19.2, 19)

plot(carbonaria.nottingham ~ year.nottingham)


year.york <- seq(from = 1990,
                 to = 2000)

carbonaria.york <- c(57.9, 61.8, 74.1, 69.6, 60, 75.0,
                     51.4, 42.1, 41.2, 37.0, 20.8)

plot(carbonaria.york ~ year.york)

year.manchester <- c("1952-1964", "1968-1974", "1990-1998")
carbonaria.manchester <- c(98.5,95.9,20.8)










year.kent <- c(1971, 1975, 1979, 1983, 1987, 
               1990, 1991, 1992, 1993, 1994)
carbonaria.kent <- c(78.2, 76.6, 71.7, 64.7, 54.3, 
                     42.3, 30.9, 33.5, 23.0, 23.8)
plot(carbonaria.kent ~ year.kent)



year.hampshire <- c("1957-1964","1965-1970","1973-1977")
carbonaria.hampshire <- c(7.1,5.8,6.5)
```



Long time series

insularia, not carboneria!

Clarke, Mani and Wynne.  1985.  Evolution in reverse: clean air and the pepperred moth.
Biological Journal of the Linnean Society 26: 189-199.
