---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Build your own dataframe part I: Vectors

## Dataframes are a type of data structure

For small datasets its often useful to build your own dataframes.  Dataframes are **objects** in R; in particular they are a type of **data structure**.  "Data structure" is just a fancy way of saying "holding or organizing data."   

We'll work with the `SeaIce` data again.
```{r}
library(Stat2Data)
data("SeaIce")
```

If we're curious about what kind of data is in an object we can use the `class()` command to see how is is classified..  Generally the first thing R spits out is the most important.

```{r}
class(SeaIce)
```

We see that `SeaIce is a `data.frame`. 

A related command is `is()`
```{r}
is(SeaIce)
```

This command is more **verbose** and tells of several things, the most relevant of which is that `SeaIce` is a dataframe.


The `SeaIce` dataframe has several columns.  We can pull up their names using the `names()` command.

```{r}
names(SeaIce)
```

There are times when we may want to examine just a single column.  We can isolate a single column using a species notation which uses a dollar sign.  To get the `Extent` column we do this

<!-- Add practice problems for this -->

```{r}
SeaIce$Extent
```


## Dataframes are made from vectors

Running this we get just a stream of numbers.  In R, such a stream of numbers is called a **vector**.  Like dataframes, vectors are a type of data structure.  We can check this with `is()`

```{r}
is(SeaIce$Extent)
```

The first result is `numeric`, which indicates that there are numbers in this vector; the second result is vector.

All columns in a dataframe are are vectors:

```{r}
is(SeaIce$Year)
is(SeaIce$Area)
```

## Make your own vectors

We can make our own vectors using the `c()` command.  (`c()` does MANY things in R, so we'll see it a lot).

Here is some more recent data on the Canda Lynx from 1984 to 2002 (Poole 2003, Table 3):
```{r}
c(13445, 8625, 6853, 6953, 6574,
  8265, 9977, 7579, 11542, 7180,
  4713, 4907, 2819, 5171, 6873, 
  6148, 8573, 9361, 11226)
```

We can save this to an R object using a very species function in R called the **assigment operator**

```{r}
lynx.ca <- c(13445, 8625, 6853, 6953, 6574,
  8265, 9977, 7579, 11542, 7180,
  4713, 4907, 2819, 5171, 6873, 
  6148, 8573, 9361, 11226)
```

We've now made a brand new object in R called `lynx.ca`.  We can see what it is by just typing its name in the console...

```{r}
lynx.ca
```
 
...and confirming what it is with `is()` and `class()`

```{r}
is(lynx.ca)
class(lynx.ca)
```


The data are from 1984 to 2002.  We can make a `year` vector like this

```{r}
year <- c(1984, 1985, 1986, 1987,
          1988, 1989, 1990, 1991, 1992, 
          1993, 1994, 1995, 1996, 1997,
          1998, 1999, 2000, 2001, 2002)
```

Its rather tedious to do that, so R has a trick.  If we want a sequence of numbers we can use the `seq()` command, which has the arguments `from =...` and `to = ...`.

```{r}
year <- seq(from = 1984, to = 2002)
```

## Checking the length of vectors

Typing in data directly into R is errorprone and whenever we do it we should check our work carefully.  A first check should be that we entered in all the numbers; we can do this by getting R to tell us the length of the vector.  Can you guess what the command is called?

```{r}
length(lynx.ca)
```

Let's check that our year vector is the same length

```{r}
length(year)
```


## Another example

Another famous example

Howlett and Majerus.  1987. The understanding of industrial melanism in the peppered moth (*Biston betularia*) .(Lepidoptera: Geometridae).  30: 31-44.

Table 2

```{r}
year.ne <- c(1954, 1961, 1970, 1975, 1981, 1982, 1983, 1984, 1985)

moths.ne <- c(92.95, 94.78, 75, 64.7, 45.9, 50, 42.9, 42.1, 39.6)
```


(Raw data is slightly more complicated, but this works)


```{r}
length(year.ne)
length(moths.ne)
```


```{r}
moths <- data.frame(year = year.ne,
                    moths.N = moths.ne)
```

```{r}
plot(moths.N ~ year, data = moths)
```

Not how R doesn't need extra space for the years where there are no data



## Try it yourself

The original lynx data that comes with R spans the year 1821–1934.  Make a sequence of numbers called year.lynx using the `seq()` command.

```{r}
1957:1964
```


There was a hypothesis that sunspot number and/or size impact lynx populations.  R has a dataset called sunspots that runs from 1749 to 1983.  Make a vector called year year.sunspot using the `seq()` command. 

```{r}
1749:1983
```


## A complicate dataframe to make

## Build the dataframe

This code makes all of the columns and makes a dataframe

```{r}
aa        <-c('A','C','D','E','F','G','H','I','K','L','M','N','P','Q','R','S','T','V','W','Y')
MW.da     <-c(89,121,133,146,165,75,155,131,146,131,149,132,115,147,174,105,119,117,204,181)
volume    <-c(67,86,91,109,135,48,118,124,135,124,124,96,90,
              114,148,73,93,105,163,141)
bulkiness <-c(11.5,13.46,11.68,13.57,19.8,3.4,13.69,21.4,
              15.71,21.4,16.25,12.28,17.43,
              14.45,14.28,9.47,15.77,21.57,21.67,18.03)
polarity  <-c(0,1.48,49.7,49.9,0.35,0,51.6,0.13,49.5,0.13,
              1.43,3.38,1.58,3.53,52,1.67,1.66,0.13,2.1,1.61)
isoelectric.pt <-c(6,5.07,2.77,3.22,5.48,5.97,7.59,6.02,9.74,5.98,
                   5.74,5.41,6.3,5.65,10.76,5.68,6.16,5.96,5.89,5.66)
hydrophobe.34 <-c(1.8,2.5,-3.5,-3.5,2.8,-0.4,-3.2,4.5,-3.9,3.8,1.9,
                  -3.5,-1.6,-3.5,-4.5,-0.8,-0.7,4.2,-0.9,-1.3)
hydrophobe.35 <-c(1.6,2,-9.2,-8.2,3.7,1,-3,3.1,-8.8,2.8,3.4,-4.8,
                  -0.2,-4.1,-12.3,0.6,1.2,2.6,1.9,-0.7)
saaH2O       <-c(113,140,151,183,218,85,194,182,211,180,204,158,
                 143,189,241,122,146,160,259,229)
faal.fold    <-c(0.74,0.91,0.62,0.62,0.88,0.72,0.78,0.88,0.52,
                 0.85,0.85,0.63,0.64,0.62,0.64,0.66,0.7,0.86,0.85,0.76)
polar.req    <-c(7,4.8,13,12.5,5,7.9,8.4,4.9,10.1,4.9,5.3,10,
                 6.6,8.6,9.1,7.5,6.6,5.6,5.2,5.4)
freq        <-c(7.8,1.1,5.19,6.72,4.39,6.77,2.03,6.95,6.32,
                10.15,2.28,4.37,4.26,3.45,5.23,6.46,5.12,7.01,1.09,3.3)
charge<-c('un','un','neg','neg','un','un','pos','un','pos',
          'un','un','un','un','un','pos','un','un','un','un','un')
hydropathy<-c('hydrophobic','hydrophobic','hydrophilic','
              hydrophilic','hydrophobic','neutral','neutral','hydrophobic','hydrophilic','hydrophobic','hydrophobic','hydrophilic','neutral',
              'hydrophilic','hydrophilic','neutral','neutral',
              'hydrophobic','hydrophobic','neutral')
volume.cat<-c('verysmall','small','small','medium',
              'verylarge','verysmall','medium','large','large',
              'large','large','small','small','medium','large','verysmall','small','medium','verylarge','verylarge')
polarity.cat<-c('nonpolar','nonpolar','polar','polar',
                'nonpolar','nonpolar','polar','nonpolar',
                'polar','nonpolar','nonpolar','polar','nonpolar','polar',
                'polar','polar','polar','nonpolar','nonpolar','polar')
chemical<-c('aliphatic','sulfur','acidic','acidic','aromatic',
            'aliphatic','basic','aliphatic','basic','aliphatic','sulfur',
            'amide','aliphatic','amide','basic','hydroxyl','hydroxyl',
            'aliphatic','aromatic','aromatic')

aa_dat <- data.frame(aa,MW.da,volume,bulkiness,
polarity,isoelectric.pt,hydrophobe.34,hydrophobe.35,
saaH2O,faal.fold, polar.req,freq,charge,hydropathy,volume.cat,polarity.cat,chemical)

```


## References

Poole, Kim  G.  2003.  A  review  of  the  Canada  Lynx, *Lynx  canadensis* ,in  Canada.  Canadian  [Field-Naturalist 117: 360-376.[(https://www.canadianfieldnaturalist.ca/index.php/cfn/article/view/738)]  DOI: https://doi.org/10.22621/cfn.v117i3.738
