# Packages & their dependencies

TODO: change from crabs to something else

R packages frequently use other R packages (which frequently use other R packages...).  When an R package requires another package, its called a **dependency.**  Depending on who and how the package was written up, dependencies might not be an issue or could be a problem.

As noted above when you download packages using RStudio's point and click interface there's a box that should be checked called "Install dependencies."

If you are using install.packages() there is an extra argument "dependencies = TRUE" that elicits the same behavior.  I'll use this to get an add-on for ggplot2 called ggpubr.

```{r eval = FALSE}
install.packages("ggpubr",dependencies = TRUE)
```

We can then install this
```{r}
library(ggpubr)
library(MASS)
```


---------------

## Optional: Make a plot with ggpubr

**This section is optional**

ggpubr is an add on to ggplot.  (This means that ggpubr has ggplot as a dependency).  Note that the syntax for ggpubr function we use, ggscatter(), has a different syntax (again) than ggplot's qplot() function and base R's plot() function.

```{r}
ggscatter(data = crabs,y = "FL", x = "RW") # use quotes!
```


**End optional section**

---------------


## Challenge

An another add-on to ggplot2 is cowplot, which stands for ["Claus O. Wilke Plot"](https://cran.r-project.org/web/packages/cowplot/vignettes/introduction.html).  Download cowplot from CRAN using either the point-and-click method or **install.packages**, and then load it using **library**. Then run the following R code, which should make the plot below.

**Note that "FL" and "RW" are NOT in quotation marks as they are for ggscatter()!**

```{r eval = FALSE}
qplot(data = crabs, y = FL, x = RW) #no quotes!
```

```{r echo = FALSE}
#load cowplot
library(cowplot)

#make the graph
qplot(data = crabs, y = FL, x = RW)
```
