---
title: 'Intro to Plotting in R: Base R graphics'
author: "brouwern@gmail.com"
date: "August 1, 2017"
output: html_document
---

```{r   include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Things to cover*
boxplot - base, qplot, ggplot
boxplot ordered
histogram 
scatterplot
scatterplot w/vert/hort/1:1
scatterplot with regression lines
scatterplot with spline smoother

scatterplot matrix


## Reference

Martin, TE.  1995.  Avian Life History Evolution in Relation to Nest Sites, Nest Predation, and Food.  Ecological Monographs.
http://onlinelibrary.wiley.com/doi/10.2307/2937160/full

**Abstract:**
"Food limitation is generally thought to underlie much of the variation in life history traits of birds. I examined variation and covariation of life history traits of 123 North American Passeriformes and Piciformes in relation to nest sites, nest predation, and foraging sites to examine the possible roles of these ecological factors in life history evolution of birds. Annual fecundity was strongly inversely related to adult survival, even when phylogenetic effects were controlled. Only a little of the variation in fecundity and survival was related to foraging sites, whereas these traits varied strongly among nest sites. Interspecific differences in nest predation were correlated with much of the variation in life history traits among nest sites, although energy trade—offs with covarying traits also may account for some variation. For example, increased nest predation is associated with a shortened nestling period and both are associated with more broods per year, but number of broods is inversely correlated with clutch size, possibly due to an energy trade—off. Number of broods was much more strongly correlated with annual fecundity and adult survival among species than was clutch size, suggesting that clutch size may not be the primary fecundity trait on which selection is acting. Ultimately, food limitation may cause trade—offs between annual fecundity and adult survival, but differences among species in fecundity and adult survival may not be explained by differences in food abundance and instead represent differing tactics for partitioning similar levels of food limitation. Variation in fecundity and adult survival is more clearly organized by nest sites and more closely correlated with nest predation; species that use nest sites with greater nest predation have shorter nestling periods and more broods,yielding higher fecundity, which in turn is associated with reduced adult survival. Fecundity also varied with migratory tendencies; short—distance migrants had more broods and greater fecundity than did neotropical migrants and residents using similar nest sites. However, migratory tendencies and habitat use were confounded, making separation of these two effects difficult. Nonetheless, the conventional view that neotropical migrants have fewer broods than residents was not supported when nest site effects were controlled."


###### Load Martin 1995 data
```{r}
martin <- read.csv("martin_1995_avian_life_history_evolution_appendix1_CSV.csv", skip = 1,
                  na.strings = c("NA","na") )
```


Change the "non-excavating" group to "secondary cavity"
```{r}
martin$group <- gsub("non-excavating","2ndary cav.",martin$group)

```


Change the "group-nesting" group to "ground"
```{r}

martin$group <- gsub("ground-nesting","ground",martin$group)
```



#Make some more changes to shorten the group names

```{r}
martin$group <- gsub("Subcanopy or canopy","subcan-/canopy",martin$group)

martin$group <- gsub("shrub or low foliage","shrub",martin$group)

martin$group <- gsub("excavating","excav",martin$group)
```



Classify general lattitudinal region as tropical, boreal, or temperate
```{r}
martin$trop.or.temp <- "temperate"
i.tropics <- which(martin$lat < 23)
i.boreal <- which(martin$lat > 54)

martin$trop.or.temp[i.tropics] <- "tropics"
martin$trop.or.temp[i.boreal] <- "boreal"
```

write.csv(martin, file = "martin_edit_CSV.csv")


Change name of martin dataframe to data
```{r}
data <- martin
```







#### R Plotting FAQ

###### Question
T1)How do I make simple tables in R?
T2)How do I make 2-way tables in R?
T3)How do I make 3-way tables in R?


#### Tables
Our object data has a column group giving the general location where each species in the dataframe nests and another columns for where the birds forage.

###### T1)How do I make simple 1-way tables in R?

We can make a simple table for where birds nest, where they forage and their general lattitudinal region like this:
```{r}
#Table for nest location
table(data$group)

#Table for forage location
table(data$foraging.site)

#Table for general lattitudal location
table(data$trop.or.temp)
```



A nice command to learn to use in R is with().  Here's what it can do when making a table;  If you were to read this out loud like you were telling R what to do, it would be somethign like this "With the dataframe called data, make a table from the group column."  This isn't that helpful in this situation, but for 2-way tables and other tasks it is nice.
```{r}
#Making a table with with()
with(data, table(group))
```


An aside on tables:
The summary() command can also give you simple 1-way tables.  However, the table() command will make a table out of whatever you give it, while summary() is more generic.  For example, the way that these particular data have been loaded in, R is treating information that is in text form, like nesting location, as text.  So  when I use summary, I get this
```{r}
summary(data$group)
```

R is very picky about differnt kinds of data.  In order for R to make a table from this column using the summary() command, you'd need to tell it that the data belong to different categories or different levels of "factor variable".  This is done with the factor() command.  here, I'll tell R that the group column is factor data and then make a table with summary()
```{r}
#Change character data to a factor
data$group <- factor(data$group)

#Now I get a table from summary()
summary(data$group)
```






###### T2)How do I make simple 2-way tables in R?
We can cross classify the nesting groups and foraging groups very easily, aka, we can make a 2-way table.

Note that you have to be explicit about where each piece of data from by putting "data$" in front of both columns
```{r}
#With nesting location in rows
table(data$group, data$foraging.site)

#With nesting location in columns 
table(data$foraging.site, data$group)
```

You can save yourself some typing by using the with() command.  If you read this outloud it would be "with the table dataframe, make a 2-way table from the group and foraging site columns"
```{r}
with(data, table(group, foraging.site))
```

###### T3)How do I make 3-way tables in R?
You can cross-classify categorical data as many ways as you'd like.  Let's make a 3-way classification based on nest type, foraging and lattidue.
```{r}
table(data$group, data$foraging.site, data$trop.or.temp)
```

This can be simplified a bit using with()
```{r}
with(data, table(group, foraging.site, trop.or.temp))
```



##### BP1)How do I make histograms in R?

The basics:
R makes histograms in a snap in R, something that I do not know of an efficient way to do in Excel.  Let's make a histogram of the different clutch sizes (number of eggs laid per nest) for the different species.  Since I've forgotten exactly what I called this column I'll first figure that out using the names() function, then make a histogram with hist().  R automatically "bins" the data and plots the frequency of observations in each bin along the y-axis.

```{r}
#Look at the names in the data dataframe
names(data)

#Make a histogram of clutch size
hist(data$clutch.sz)
```

Finesse: Changing titles
R automatically assigns names to everything.  Here I change the x-axis (called the x label) and title, which is called the "main title".
```{r}
hist(data$clutch.sz, xlab = "Clutch size", main = "data's clutch-size data")
```

Finesse: Adding a vertical line
The bird I study typically lays 5 eggs.  To make it easy to compare my bird to other birds, I can add a vertical line at 5.  This requires two seperate commands.  First I'll make the plot, then add the line.  The line is made with the abline() command, which stands for "A-B line", where A stands for the intercept of a line B stands for its slope.  abline() can also make horizontal lines and vertical lines, but the programmers must have thought that abhvline() was too long of a name. 
```{r}
#Make histogram
hist(data$clutch.sz)

#Add vertical line at 5
abline(v = 5)
```

The color is black so it blends in.  I can change the color using the col= comamnd within abline line
```{r}
#Make histogram
hist(data$clutch.sz)
abline(v = 5, col = 2)
```


##### BP2)How do I make boxplot in R?
Oneliner:
boxplot(data$clutch.sz ~ data$group)

Skelp
```{r}
par(mfrow = c(1,1))
boxplot(clutch.sz ~ group, data = data,
        varwidth = TRUE, #modify the width proportional to sample size
        notch = FALSE,   #notches can be used to judge if medians are          
                         #different.  Work best with relativley large sample                 
                         #sizes.
        col = 2:7)       #colors

```


Details:
Boxplots are recommended by many statisticians as an excellent way to summarize the distribution of data.  Boxplots also a breeze in R, while they are more or less impossible in Excel (I've seen instructions for them but it looks like a pain).  Let's make a box plot of clutch size compared to birds from different nesting locations.  The boxplot() function uses a tilda, ~ in it.  Like many R functions, such as those for regression, the continuous variable goes to the left of ~ and the categorical variables go to the right.
```{r}
boxplot(data$clutch.sz ~ data$group)
```




##### BP3)How do I make scatterplot in R?

*Oneliner:*
plot(clutch.sz ~ nest.dur, data = data)

*Skelp:*
```{r}
#Plot
plot(clutch.sz ~ nest.dur, data = data,
     main = "Main Title: Clutch size vs. nesting duration",
     sub = "Subtitle: Martin 199x Ecological Monographs",
     xlab = "xlab: nesting duration",
     ylab = "ylab: clutch size",
     col = 2,   #change the color
     pch = 2)   #change th shape of the symbol

#Regression line
abline(a = 3.6,   #intercept
       b = 0.045) #slope
```




*Details:*
Scatter plots are more or less the default plot type of R.  There are several minor variation that all do the same thing.  The 1st is probably the easiest to remember because it is similar to the standard format for many other R commands, such as linear regression.
```{r}
#Using a tilda ~ and "data =""
plot(clutch.sz ~ nest.dur, data = data)

#Using tilda and explicit column references
plot(data$clutch.sz ~ data$nest.dur)

#Using commas and explicit column ref.  NOte that order of data has been flipped.
plot(data$nest.dur, data$clutch.sz)

```



##### BP4)How do I put a regression line through a  scatterplot in R?

Oneliner:
plot(clutch.sz ~ nest.dur, data = data)
abline(a = 3.6, b = 0.045)

Details

```{r}

```



##### BP5)How do I put 2 plots next to each other in R?

*Oneliner:*
par(mfrow = c(1,2))
plot(clutch.sz ~ nest.dur, data = data)
plot(clutch.sz ~ surv.adult, data = data)


*Details*
R can relativley easily format plots next to each other, but the function is rather cryptic, involving the par() command with "mfrow =" in.  To plot two figures next to each other use 'par(mfrow = c(1,2))', which basically says "put two plots in a 1 by 2 grid."  The default setting for R is par(mfrow = c(1,1)), which corresponds to a 1 x 1 grid.  You can probably guess that to put two plots on top of each other you would use par(mfrow = c(2,1)), and so forth.

```{r eval - F}
#Two plots next to each other
par(mfrow = c(1,2))
plot(clutch.sz ~ nest.dur, data = data)
plot(clutch.sz ~ surv.adult, data = data)


#Two plots on top of each other
par(mfrow = c(2,1))
plot(clutch.sz ~ nest.dur, data = data)
plot(clutch.sz ~ surv.adult, data = data)

```



*Finnese:*
Lots of tweaks can be done to plot in general, and to multiple plots in a grid, such as the space between each plots.  The code, however, is rather cryptic.
If you look at the help menu for the par() command you'll see dozen of options.  For quickly making plots look nice, many people just export into Power Point and orient and annotat them there.  Learning how to feed commands to par() to make plots and groups of plots look nice can save you time in the long run, but it takes time to figure out what you want and what you like.  I find that the R package ggplot2 with its function qplot() makes nicer default plots.  GGPLOT has its own complicated syntax, but is prefferred by many R users.




##### How do I put two plots next to each other using ggplot?
Oneliner
library(ggplot2)
qplot(y = pred, x = nest.dur, data = data)


##### How do I add a regression line to a scatterplot with ggplot?
Oneliner
library(ggplot2)
qplot(y = pred, x = nest.dur, data = data,
geom = c("smooth","point"), method = "lm")


Skelp
qplot(y = pred, x = nest.dur, data = data,
geom = c("smooth","point"),
se = FALSE,
method = "lm",
xlab = "xlab = nesting duration",
ylab = "ylab = predation rate",
main = "main title: predation vs. nesting duration")






#### How do I make a strip chart
Box plot are great but don't work well for small data set.

```{r}
#Vertical strip chart
stripchart(data$clutch.sz ~ factor(data$group),
           vertical = TRUE)

#Horizontal strip chart
stripchart(data$clutch.sz ~ factor(data$group),
           vertical = FALSE)
```


