# A guide to current swirl tutorials 

Below are the names of the swirl tutorials which are most relevant to new R users, how to download them and the names of the units within each tutorial 

## Regression Models: The basics of regression modeling in R (Team swirl)

Installation:
```{r, eval = F}
swirl::install_course("Regression Models")
```

**Units:**

1. Introduction
1. Residuals
1. Least Squares Estimation
1. Residual Variation
1. Introduction to Multivariable Regression
1. MultiVar Examples
1. MultiVar Examples2
1. MultiVar Examples3
1. Residuals Diagnostics and Variation
1. Variance Inflation Factors
1. Overfitting and Underfitting
1. Binary Outcomes
1. Count Outcomes


## Statistical Inference: The basics of statistical inference in R (Team swirl)

Installation:
```{r, eval = F}
swirl::install_course("Statistical Inference")
```

**Units:**

1. Introduction
1. Probability1
1. Probability2
1. ConditionalProbability
1. Expectations
1. Variance                
1. **CommonDistros**           
1. Asymptotics
1. **T Confidence Intervals**
1. **Hypothesis Testing**     
1. **P Values**               
1. Power
1. **Multiple Testing**       
1. **Resampling**          


## Exploratory Data Analysis: The basics of exploring data in R (Team swirl)

Installation
```{r, eval = F}
swirl::install_course("Exploratory Data Analysis")
```


1. Principles of Analytic Graphs
1. Exploratory Graphs
1. Graphics Devices in R
1. Plotting Systems
1. Base Plotting System
1. Lattice Plotting System
1. Working with Colors
1. GGPlot2 Part1
1. GGPlot2 Part2
1. GGPlot2 Extras
1. Hierarchical Clustering
1. K Means Clustering
1. Dimension Reduction
1. Clustering Example
1. CaseStudy

## Getting and Cleaning Data (Team swirl)

Installation
```{r, eval = F}
swirl::install_course("Getting and Cleaning Data")
```

**Units**
1: Manipulating Data with dplyr
1: Grouping and Chaining with dplyr
1: Tidying Data with tidyr
1: Dates and Times with lubridate

## Advanced R Programming (Roger Peng)

```{r, eval = F}
swirl::install_course("Advanced R Programming")

```


1. Setting Up Swirl
1. Functions
1. Functional Programming with purrr0

## The R Programming Environment (Roger Peng)

```{r, eval = F}
swirl::install_course("The R Programming Environment")

```

1. Setting Up Swirl
1. Basic Building Blocks
1. Sequences of Numbers
1. Vectors
1. Missing Values
1. Subsetting Vectors
1. Matrices and Data Frames
1. Logic
1. Workspace and Files
1. Reading Tabular Data
1. Looking at Data
1. Data Manipulation
1. Text Manipulation Functions
1. Regular Expressions
1.  The stringr Package

## Regular Expressions (Jon Calder)

Installation:
```{r, eval = F}
swirl::install_course("Regular Expressions")
```

**Units**

1. Introduction
1. Regex in base R
1. Anchors
1. Character Classes
1. Groups and Ranges
1. Quantifiers
1. Applied Examples

## A (very) short introduction to R (Claudia Brauer)

```{r, eval = F}
swirl::install_course("A_(very)_short_introduction_to_R")
```

The author notes: 
>"Thiscourse is based on a non-interactive tutorial with the same name, which can be downloaded from www.github.com/ClaudiaBrauer/A-very-short-introduction-to-R. The contents are the same (with a few exceptions), so you can open the pdf version alongside to look up how to do something you learned before or browse through the references on the last two pages."

This tutorial has 3 modules.  The first is introductory.  The others get into more details tasks, such as sourceing .R file scripts.



<!-- Data Science and R -->
<!-- source("http://wush978.github.io/R/init-swirl.R") -->

<!-- Google Forms Course -->
<!-- Sean Cross -->
<!-- "An example course that instructors can use as a template for collecting students’ progress through a swirl course. This approach is recommended for courses that are taught in person." -->
<!-- swirl::install_course_github("seankross", "Google_Forms_Course") -->


## R Programming: The basics of programming in R (team swirl)


1. Basic Building Blocks      
1. Workspace and Files
1. Sequences of Numbers       
1. Vectors
1. Missing Values             
1. Subsetting Vectors
1. Matrices and Data Frames   
1. Logic
1. Functions                 
1. lapply and sapply
1. vapply and tapply        
1. Looking at Data
1. Simulation                
1. Dates and Times
1. Base Graphics


The key vocabulary, concepts and functions  covered in this tutorial are:

* programming language
* variable
* `+`, `-`, `/`, and `^`
* sqrt()
* abs()
* assignment operator
* <-
* help files
* vector
* c() function ("concatenate", "combine")
* element wise operation ("element-by-element")
* recycling
* vectorized operations (not discussed in those terms)
* up arrow to view command history
* tab completion ("auto-completion")

```{r, echo = F, eval = F}
# all commands used
5 + 7
x <- 5+7
x
x-3
y <- x-3
z<-  c(1.1, 9, 3.14)
?c
z
c(z, 555,z)
z*2+100
my_sqrt <- sqrt(z-1) # results in vector
my_sqrt
my_div <- z/my_sqrt
c(1,2,3,4) + c(0,10) # vector recycling
c(1, 2, 3, 4) + c(0, 10, 100)
z*2+1000 # via up arrow
```





