# Improving bas R plots: moving axes labels with mtext()

Some notes on adjusting margins and using the mtext() function to place axis labels closer the axes.

# Standard plot 1 panel plot

## Issues

* Large margin above the plot
* Large distance between labels on the tick marks and the axis label

```{r}
data(iris)
with(iris, plot(Sepal.Length ~ Petal.Length))
```



## Change margins around plot

* use "par(mar = c(lower, left,top,right))" 
* defult is mar =  c(5, 4, 4, 2) + 0.1.

Here, I decrease the top and left margins to 1.
```{r}
par(mar = c(5,4,1,1))
with(iris, plot(Sepal.Length ~ Petal.Length))
```




## Move axis labels closer to axis

* set axis labels to xlab = "" and ylab = ""
* Add new labels using mtext()
* mtext stands for "margin text"
* mtext takes several arguements:

Here, I decrease the top and left margins to 1.



### Change axis lables
```{r}
with(iris, plot(Sepal.Length ~ Petal.Length,
      xlab = "Petal Length",
      ylab = "Sepal length"))
```


### Plot with no axis labels

* ylab = "" sets it to print nothing
```{r}
with(iris, plot(Sepal.Length ~ Petal.Length,
      xlab = "",
      ylab = ""))
```


### Plot with x axis added with mtext()

* side = 1 sets it to the lower maring
* side = 2 would do the left
* side= 3 top, side = 4 right
* line = 2 sets it 2 lines below the axis
* it usually takes some trial and error to get it where you want it
```{r}
with(iris, plot(Sepal.Length ~ Petal.Length,
      xlab = "",
      ylab = ""))

#add text 
mtext(text = "Petal Length",
      side = 1,#side 1 = bottom
      line = 2)
```




### Plot with y axis added with mtext()

* side = 2 would do the left

```{r}
with(iris, plot(Sepal.Length ~ Petal.Length,
      xlab = "",
      ylab = ""))

# x axis
mtext(text = "Petal Length",
      side = 1,line = 2)

# y axis

mtext(text = "Sepal Length",
      side = 2, #side 2 = left
      line = 2)
```


### Reset margins 

* Get rid of blank space by changing margins
* set mar = c(2,2,1,1)

```{r}

par(mar = c(3,3,1,1))

with(iris, plot(Sepal.Length ~ Petal.Length,
      xlab = "",
      ylab = ""))

# x axis
mtext(text = "Petal Length",
      side = 1,line = 2)

# y axis

mtext(text = "Sepal Length",
      side = 2, #side 2 = left
      line = 2)
```


# No tick marks & tick labels

If for some reason you want no tick marks and no numbers along the axies, you can set par(yaxt = "m")

```{r}
#change par
par(yaxt="n",#remove from y
    xaxt='n')#remove from x


with(iris, plot(Sepal.Length ~ Petal.Length))
```


Set it back tiwht yaxt = "t"
```{r}
#change par
par(yaxt="t",#remove from y
    xaxt='t')#remove from x


with(iris, plot(Sepal.Length ~ Petal.Length))
```

