# R's Installation directory and "working directory"

When using R there are several levels of where software, files, and R's working memory are located and organized.

First, R is a program that lives on your hard drive.  Occasionally it may get installed someplace we don't want.  We can check where R is located using the `R.home()` function.  This isn't a commonly used function but is useful for troubleshooting problems.

Second, when R is running it has a **working directory** where any files you save will be stored.  You can check where your working directory is using `get.wd()`.    When you get started, you don't have to worry much about this, but eventually, it will be *very* important to make sure you know where your files are going so you can find them back!  Using **R Projects** managed with RStudio is very useful when the time comes to do this.

Third, often in R we create **objects** (sometimes called **variables**) which contain information.  These are made using R's assignment operator: `<-`, which is a "less than" sign and a dash (or an alligator eating a jumping fish, whichever you prefer).  To assign something to R's working memory I do this

```{r}
my.object <- "something"
```


I can see everything I've assigned to R's working memory using the command `ls()`.
```{r}
ls()
```




This video covers similar topics

```{r, echo = F}
knitr::include_url("https://youtu.be/jRT8zOiQtsE")
```

The code used in the video is below.  You can just copy and paste it into your R console if you have any trouble.

```{r, eval = F}
my.name <- "Your name" #character string
is(my.name)
my.name
home <- R.home()
home
wd <- getwd()
wd
ls()
```

