# (PART) Getting Software & Data Into R {-}

### {-}



# Loading packages from CRAN

```{r include=FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE,fig.height = 3, fig.width = 5)
```




## Introduction



### Learning objectives

This exercise will introduce students to

1. the concept of an **R Package** (aka **library**)
1. how to load R packages using the library() function
1. the R plotting package **ggplot2**
1. cool add-ons to ggplot2, **ggpubr** and **cowplot**

### Learning goals

By the end of this exercise students should be able to 

* locate and download packages from the CRAN website using RStudio
* recognize the R functions used to download and install packages.




### Packages 

* MASS
* ggplot2
* ggpubr
* cowplot

### Potential hangups

* quoted vs. unquoted text (eg qplot vs. ggpubr syntax)


## Loading packages that come with base R

What you download from CRAN is **base R** (also known as the **base distribution**).  Many functions are called **base functions** because they are hard-wired into R.  




### Loading packages with the library() function

When a function is already downloaded to your computer, you explicitly load it into R's working memory using the library() command.

```{r}
library(MASS)
```
















