---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Summary stats for main PA eagle dataframe

```{r}
library(compbio4all)
data("eaglesPA")
```


There are many commands for summary data in R, such as mean, median.  However, you have to be careful about NAs!

```{r}
mean(eaglesPA$eagles)
```

So, you asked for the mean of the eagles data, and you got NA.  That's really annoying.  


### "NA" is a big deal in R

Try this
```{r}
mean(eaglesPA$eagles, na.rm = T)

```
"na.rm = T", which means "na.rm = TRUE", which means, "should I remove the NAs = yes, do it"


### R also is VERY picky about uppper vs. lower case

```{r, error = T}
Mean(eaglesPA$eagles, na.rm = T)

```

Note that the R error message is not very helpful : (



### Most basic R commands are lower case

```{r, eval = F}
mean(eaglesPA$eagles, na.rm = T)
median(eaglesPA$eagles, na.rm = T)
min(eaglesPA$eagles, na.rm = T)
max(eaglesPA$eagles, na.rm = T)
summary(eaglesPA$eagles)
sd(eaglesPA$eagles, na.rm = T)

```

### Standard error in R

The standard error (se) is a very common summary statistics but for some reason 
there is not a function for it in base R

#### Calcualte the standard error by hand

Use the sd() command and the square root command sqrt
```{r}
sd(eaglesPA$eagles, na.rm = T)/sqrt(15)
```
