---
output: html_document
editor_options: 
  chunk_output_type: console
---


## Doing math for stats by hand

```{r, echo = T}
library(compbio4all)
data(genes_hpc)
```

Put gene lengths in vector

```{r}
gene_lengths <- genes_hpc$length
```


Make  subset so that things can all be seeing on-screen
Then do for whole vector
Assignment: do for GC
Follow up: data manipulation with dplyr: 
1)2 groups: mt vs rest of genome
2)3 groups: mt vs. regular vs. sex
3)may groups: by chromosome number



### Calculating the mean gene length by hand

#### The numerator
```{r}
length.sum <- sum(gene_lengths)
```

### The denomintor
```{r}
N <- length(gene_lengths)

```

### The mean
```{r}
length_mean <- length.sum/N


```


Same thing to get the mean, just in 1 step
```{r}
 sum(gene_lengths)/length(gene_lengths)

```




We can check our answer like this using the "==" function which asks R "are these two objects EXACTLy the same?"

```{r}
length_mean == mean(gene_lengths)
```




We can do the same thing to check that the standard deviation is indeed the square root of the mean

```{r}
sd(gene_lengths) == sqrt(var(gene_lengths))
```

If the were not the same, R would say "FALSE"


* Because R is very very very very precise any rounding errors will result in R saying that two things are NOT the same.  Sometimes some R functions do do some rounding so you have to check "FALSE" sometimes.



